\begin{thebibliography}{9}

\bibitem{ourworldindata}
C. Giattino, E. Mathieu, V. Samborska, and M. Roser, ``Data Page: Exponential growth of computation in the training of notable AI systems,'' 
Our World in Data, 2023. [Online]. Available: \href{https://ourworldindata.org/grapher/exponential-growth-of-computation-in-the-training-of-notable-ai-systems}{ourworldindata.org} (accessed: Oct. 2025).

\bibitem{semico2023riscv}
Semico Research Corp., ``RISC-V CPU Market to Exceed 10B Cores by 2030,'' 
Market Report, Semico Research, 2023.

\bibitem{riscv-v-spec}
RISC-V International, ``The RISC-V Vector ISA Extension, Version 1.0,'' 2021. 
[Online]. Available: \href{https://github.com/riscv/riscv-v-spec}{GitHub: riscv-v-spec} (accessed: Jul. 2025).

\bibitem{onnx}
ONNX Project, ``Open Neural Network Exchange (ONNX),'' 2025. 
[Online]. Available: \href{https://onnx.ai/}{onnx.ai} (accessed: Oct. 2025).

\bibitem{vicuna}
Michael Platzer and Peter Puschner. Vicuna: A Timing-Predictable RISC-V Vector Coprocessor for Scalable Parallel Computation. 
In \textit{33rd Euromicro Conference on Real-Time Systems (ECRTS 2021)}. Leibniz International Proceedings in Informatics (LIPIcs), Volume 196, pp. 1:1-1:18,
Schloss Dagstuhl – Leibniz-Zentrum für Informatik (2021) \url{https://doi.org/10.4230/LIPIcs.ECRTS.2021.1}

\bibitem{ara}
Matteo Perotti, Matheus Cavalcante, Nils Wistoff, Renzo Andri, Lukas Cavigelli, and Luca Benini. A "New Ara" for Vector Computing: An Open Source Highly
Efficient RISC-V V 1.0 Vector Processor Design. In \textit{2022 IEEE 33rd International Conference on Application-specific Systems, Architectures and 
Processors (ASAP)}, 2022. \url{https://doi.org/10.1109/ASAP54787.2022.00017}

\bibitem{riscv-gnu-toolchain}
RISC-V International Collaboration, ``RISC-V GNU Compiler Toolchain.'' 
[Online]. Available: \href{https://github.com/riscv-collab/riscv-gnu-toolchain}{GitHub: riscv-gnu-toolchain} (accessed: Jun. 2025).

\bibitem{qemu-riscv}
QEMU Project, ``QEMU: Open Source Machine Emulator and Virtualizer (RISC-V Target).'' 
[Online]. Available: \href{https://www.qemu.org/}{qemu.org} (accessed: Jun. 2025).

\bibitem{verilator}
W. Snyder and contributors, ``Verilator: Open-Source SystemVerilog Simulator.'' 
[Online]. Available: \href{https://www.veripool.org/verilator/}{veripool.org/verilator} (accessed: Jul. 2025).

%%%%%%%%%%%%%%%%%%%%%%%%%
% Papers
%%%%%%%%%%%%%%%%%%%%%%%%%

\bibitem{herdt2018extensible}
V. Herdt, D. Große, H. M. Le and R. Drechsler, "Extensible and Configurable RISC-V Based Virtual Prototype," in \emph{2018 Forum on Specification \& Design Languages (FDL)}, Garching, Germany, 2018, pp. 5--16, doi: 10.1109/FDL.2018.8524047.

\bibitem{schlaegl2024riscv}
M. Schlägl, M. Stockinger and D. Große, "A RISC-V “V” VP: Unlocking Vector Processing for Evaluation at the System Level," in \emph{2024 Design, Automation \& Test in Europe Conference \& Exhibition (DATE)}, Valencia, Spain, 2024, pp. 1--6, doi: 10.23919/DATE58400.2024.10546838.

\bibitem{quiroga2022reusable}
J. Quiroga, R. I. Genovese, I. Diaz, H. Yano, A. Ali, N. Sonmez, O. Palomar, V. Jimenez, M. Rodriguez, and M. Dominguez, "Reusable Verification Environment for a RISC-V Vector Accelerator," in \emph{Proceedings of the Design and Verification Conference \& Exhibition Europe (DVCon Europe)}, Munich, Germany, Dec. 2022, pp. 1--8.

\bibitem{imianosky2024reliability}
C. Imianosky, D. A. Santos, D. R. Melo, F. Viel and L. Dilillo, "Special Session: Reliability and Performance Evaluation of a RISC-V Vector Extension Unit for Vector Multiplication," in \emph{2024 IEEE International Symposium on Defect and Fault Tolerance in VLSI and Nanotechnology Systems (DFT)}, Didcot, United Kingdom, 2024, pp. 1--6, doi: 10.1109/DFT63277.2024.10753524.

\bibitem{alassir2021arrow}
I. Al-Assir, M. M. Yildirim, and U. Y. Ogras, "Arrow: A RISC-V Vector Accelerator for Machine Learning Inference," in \emph{2021 IEEE 32nd International Conference on Application-specific Systems, Architectures and Processors (ASAP)}, 2021, pp. 123--130.

\bibitem{wang2024speed}
Y. Wang, C. Zhang, Z. Wang, and H. Li, "A Scalable RISC-V Vector Processor Enabling Efficient Multi-Precision DNN Inference," in \emph{IEEE Transactions on Very Large Scale Integration (VLSI) Systems}, 2024.

\bibitem{carpentieri2025performance}
L. Carpentieri, M. VazirPanah and B. Cosenza, "A Performance Analysis of Autovectorization on RVV RISC-V Boards," in \emph{2025 33rd Euromicro International Conference on Parallel, Distributed, and Network-Based Processing (PDP)}, Turin, Italy, 2025, pp. 129--136, doi: 10.1109/PDP66500.2025.00026.

\bibitem{volokitin2023case}
V. Volokitin, E. Kozinov, V. Kustikova, A. Liniov, and I. Meyerov, "Case Study for Running Memory-Bound Kernels on RISC-V CPUs," \emph{arXiv preprint} arXiv:2305.09266, 2023, doi: 10.48550/arXiv.2305.09266.

\bibitem{brown2025riscv}
N. Brown, "Is RISC-V ready for High Performance Computing? An evaluation of the Sophon SG2044," \emph{arXiv preprint} arXiv:2508.13840, submitted Aug. 2025, doi: 10.48550/arXiv.2508.13840.

%%%%%%%%%%%%%%%%%%%%%%%%%

\bibitem{lecun1998gradient}
Y. Lecun, L. Bottou, Y. Bengio, and P. Haffner, ``Gradient-based learning applied to document recognition,'' \emph{Proceedings of the IEEE}, vol. 86, no. 11, pp. 2278--2324, 1998, doi: 10.1109/5.726791.

\bibitem{redmon2016yolo}
J. Redmon and A. Farhadi, ``YOLO9000: Better, Faster, Stronger,'' \emph{arXiv preprint} arXiv:1612.08242, 2016, doi: 10.48550/arXiv.1612.08242.

\bibitem{spike-simulator}
RISC-V International, ``Spike: RISC-V ISA Simulator,'' 2024. 
[Online]. Available: \href{https://github.com/riscv-software-src/riscv-isa-sim}{GitHub: riscv-isa-sim} (accessed: Jul. 2025).

\bibitem{cv32e40x}
OpenHW Group, ``CV32E40X RISC-V Core.'' 
[Online]. Available: \href{https://github.com/openhwgroup/cv32e40x}{GitHub: cv32e40x} (accessed: Jul. 2025).

\bibitem{cva6}
F. Zaruba and L. Benini, ``The Cost of Application-Class Processing: Energy and Performance Analysis of a Linux-Ready 1.7-GHz 64-Bit RISC-V Core in 22-nm FDSOI Technology,'' \textit{IEEE Transactions on Very Large Scale Integration (VLSI) Systems}, vol. 27, no. 11, pp. 2629--2640, Nov. 2019.

\end{thebibliography}